### 2023-03-20 OMG_2023-03-07/OMG_2023-03-07.Rmd
big commit

---
title: "Don't Mis-Measure Measurement Model Measurement Error"
subtitle: "OMG Workshop"
author: "Bertrand Wilden"
date: \today
output: binb::metropolis
header-includes:
   - \usepackage{tikz}
   - \usetikzlibrary{positioning}
   - \usepackage{xcolor}
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache.lazy = FALSE)

pacman::p_load(
  "tidyverse",
  "here",
  "patchwork",
  "targets",
  "kableExtra",
  "tidybayes",
  "ggdist",
  "MetBrewer",
  "targets"
)

options(scipen = 1, digits = 3)
```

##

**Measurement Models**

- Used to measure latent variables: *ideology, corruption, democracy, racial resentment*

**Theory-Testing Models**

- Testing a causal theory: *being an ideologically extreme candidate hurts one's reelection chances*

##

**Using Measurement Model Outputs in Theory-Testing Models**

::: incremental
  - Because measurement models are statistical constructs they contain uncertainty estimates
  - In practice, almost everyone disregards this uncertainty and uses only the *maximum a posteriori* (MAP) values from measurement models for their theory-testing models
    - Mean, median, mode estimates from the measurement model
:::

##

**Three Problems With Ignoring Measurement Model Uncertainty**

::: incremental
  1. Uncertainty is not properly propagated in theory-testing models using MAP values
    - ***Coefficient uncertainty is underestimated***
  2. There is too much random noise in the MAP values
    - ***Coefficients are biased towards zero***
  3. Possibility of backdoor confounding through the measurement process
    - ***Coefficients are biased in unpredictable ways***
:::

##

**Measurement Model Cases**

::: incremental
  - Bayesian Improved Surname Geocoding (**BSIG**)
    - Used to measure race/ethnicity based on names and location. Measurement model is categorical
  - Item-Response Theory Ideal Points (**IRT**)
    - Used to infer ideology of political actors. Measurement model is continuous
:::

##

```{r}
set.seed(1111)

p1 <- tibble(race = c("AAPI", "Black", "Hispanic", "Native\nAmerican", "White"),
       probs = round(c(MCMCpack::rdirichlet(1, c(1, 5, 15, 3, 4))), 3)) |> 
  ggplot(aes(x = race, fill = probs, y = 0, label = probs)) +
  ggtext::geom_textbox(color = "black", height = .15, width = .15, valign = .5, halign = .5,
               size = 6) +
  geom_text(aes(label = race, y = 0.25), size = 5) +
  geom_segment(aes(x = "Hispanic", xend = "Hispanic", y = -.15, yend = -.7),
               arrow = arrow(length = unit(.1, "inches")), color = "black", linewidth = .5) +
  geom_text(aes(label = "Hispanic", x = "Hispanic", y = -.75), size = 8, color = "black") +
  geomtextpath::geom_textcurve(aes(x = 0, xend = 0, yend = 0.9, y = -.4), 
                               label = "Measurement Model", position = position_dodge(width = 1), 
                               curvature = -.1, size = 7) +
  geomtextpath::geom_textcurve(aes(x = 0, xend = 0, yend = -.5, y = -.9), 
                               label = "MAP", position = position_dodge(width = 1), 
                               curvature = -.1, size = 7) +
  theme_void() +
  ylim(-.9, 0.9) +
  scale_fill_continuous(low = "white", high = met.brewer("Isfahan1", 2)[2]) +
  theme(legend.position = "none", plot.title = element_text(hjust = .5, size = 24),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA")) +
  labs(title = "BISG")

p2 <- tibble(x = rnorm(500, mean = .75, sd = .75)) |> 
  ggplot(aes(x = x, label = round(mean(x), 3))) +
  geom_histogram(bins = 75, color = "white", fill = met.brewer("Isfahan1", 2)[2]) +
  geom_segment(aes(x = mean(x), xend = mean(x), y = 0, yend = -18), color = "white") +
  geom_segment(aes(x = mean(x), xend = mean(x), y = 3, yend = -10), 
               arrow = arrow(length = unit(.1, "inches")), color = "black", linewidth = .5) +
  geom_segment(aes(x = -5, xend = 5, y = 0, yend = 0), 
               color = "black", linewidth = 1) +
  geom_curve(aes(x = 5, xend = 5, yend = 40, y = -3),
             curvature = .15, position = position_dodge(width = 1)) +
  geom_curve(aes(x = 5, xend = 5, yend = -6, y = -18),
             curvature = .15, position = position_dodge(width = 1)) +
  geom_text(aes(x = mean(x), y = -12), size = 8, color = "black") +
  geom_text(aes(x = -3.5, y = -2), label = "Liberal", size = 6) +
  geom_text(aes(x = 3.5, y = -2), label = "Conservative", size = 6) +
  theme_void() +
  theme(plot.title = element_text(hjust = .5, size = 24),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA")) +
  labs(title = "Bayesian IRT")

p1 + p2
```

# Measurement Error and Backdoor Confounding: BSIG Example

##

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{Large}
\begin{center}
\tikz{
    \node (race_s) {$R^*$};
    \node (race_s_label) [right = .1 of race_s] {$\textcolor{teal}{\text{\normalsize Estimated Race}}$};
    \node[hidden, dashed] (race) [above = 1.5 of race_s] {$R$};
    \node (race_label) [above = .1 of race] {$\textcolor{teal}{\text{\normalsize True Race}}$};
    \node (e_race) [left = 1.5 of race_s] {$e^R$};
    \node (e_race_label) [left = 0 of e_race] {$\textcolor{teal}{\text{\normalsize Measurement Error}}$};

    \path[dashed] (race) edge (race_s);
    \path (e_race) edge (race_s);
}
\end{center}
\end{Large}

##

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{Large}
\begin{center}
\tikz{
    \node (race_s) {$R^*$};
    \node[hidden, dashed] (race) [above = 1.5 of race_s] {$R$};
    \node (e_race) [left = 1.5 of race_s] {$e^R$};
    \node[hidden, dashed] (confound) [below = 1.5 of race_s] {$U$};
    \node (confound_label) [below = .1 of confound] {$\textcolor{teal}{\text{\normalsize Confound}}$};
    \node (y) [right = 1.5 of race_s] {$Y$};
    \node (y_label) [above = .1 of y] {$\textcolor{teal}{\text{\normalsize Turnout}}$};
    \node (t) [below right = 1.5 of y] {$T$};
    \node (t_label) [below = 0 of t] {$\textcolor{teal}{\text{\normalsize Time}}$};

    \path[dashed] (race) edge (race_s);
    \path (e_race) edge (race_s);
    \path[dashed] (confound) edge [bend left=20] (e_race);
    \path[dashed] (confound) edge [bend right=20] (y);
    \path (race_s) edge (y);
    \path (t) edge (y);
}
\end{center}
\end{Large}

##

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{Large}
\begin{center}
\tikz{
    \node (race_s) {$R^*$};
    \node[hidden, dashed] (race) [above = 1.5 of race_s] {$R$};
    \node (e_race) [left = 1.5 of race_s] {$e^R$};
    \node[hidden, dashed, red] (confound) [below = 1.5 of race_s] {$U$};
    \node (y) [right = 1.5 of race_s] {$Y$};
    \node (t) [below right = 1.5 of y] {$T$};

    \path[dashed] (race) edge (race_s);
    \path (e_race) edge (race_s);
    \path[dashed, red] (confound) edge [bend left=20] (e_race);
    \path[dashed, red] (confound) edge [bend right=20] (y);
    \path (race_s) edge (y);
    \path (t) edge (y);
}
\end{center}
\end{Large}

##

\centering

\textbf{Solution: Joint Bayesian Model}

\raggedright

\begin{equation*}
\begin{aligned}
   y_i  &\sim f(\pi_i) \\
   \pi^{*}_i &\sim g(\pi_i)
\end{aligned}
\end{equation*}

- $y$, outcome in theory-testing model
- $\pi$, latent variable
- $\pi^*$, MAP estimate of latent variable
- $f(\cdot)$, theory-testing model
- $g(\cdot)$, measurement model

##

**Simplifications**

::: incremental
- Applied researchers know a lot about $f(\cdot)$, but not $g(\cdot)$
- The real $g(\cdot)$ can take a lot of computation time
- My solution is to use an approximation of $g(\cdot)$ that fits the measurement error posterior distribution
:::

# Measurement Error and Attenuation Bias: IRT Ideal Point Example

##

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{Large}
\begin{center}
\tikz{
    \node (theta_s) {$\theta^*$};
    \node (theta_s_label) [below = 0 of theta_s] {$\textcolor{teal}{\text{\normalsize Estimated Ideal Point}}$};
    \node[hidden, dashed] (theta) [above = 1.5 of theta_s] {$\theta$};
    \node (theta_label) [above = 0 of theta] {$\textcolor{teal}{\text{\normalsize True Ideal Point}}$};
    \node (e_theta) [left = 1.5 of theta_s] {$e^\theta$};
    \node (p) [below = 1.5 of e_theta] {$P$};
    \node (p_label) [below = 0 of p] {$\textcolor{teal}{\text{\normalsize Participation}}$};
    \node (y) [right = 1.5 of theta_s] {$Y$};
    \node (y_label) [above = 0 of y] {$\textcolor{teal}{\text{\normalsize Reelection}}$};

    \path (e_theta) edge (theta_s);
    \path (theta_s) edge (y);
    \path[dashed] (theta) edge (theta_s);
    \path (p) edge [bend left=0] (e_theta);
    \path[dashed] (theta) edge [bend right=30] (e_theta);
}
\end{center}
\end{Large}

##

```{r}
source(here::here("R", "sim_funcs.R"))

set.seed(100)

sim_cont_data(200, noise_x = 1, noise_a = .1, true_b = 2)$dat |> 
  arrange(x_meas) |> 
  mutate(id = row_number()) |> 
  ggplot(aes(x = x_meas, y = id)) +
  geom_pointrange(aes(xmin = x_meas - x_sd, xmax = x_meas + x_sd), 
                  alpha = .75, size = .5, color = met.brewer("Isfahan1", 1)) +
  geom_pointrange(aes(xmin = x_meas - x_sd * 1.96, xmax = x_meas + x_sd * 1.96), 
                  alpha = .5, size = .5, color = met.brewer("Isfahan1", 1)) +
  # geom_point(color = "red", size = .3, alpha = .2) + 
  labs(x = expression(paste("Estimated Ideal Point, ", theta^{`*`})), y = "Groups (sorted)",
       caption = "Means and Standard Errors", title = "Simulated Random Normal Error") +
  theme_ggdist() +
  theme(text = element_text(family = ""),
        plot.title = element_text(size = 22),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        plot.caption = element_text(size = 18),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"))
```

##

**Joint Model for Normal Measurement Error**

\begin{equation*}
\begin{aligned}
  y_i &\sim \text{Normal}(\mu_i, \sigma) \\[-1pt]
  \mu_i &= \beta_0 + \beta_1 X_{\text{TRUE,i}} \\[-1pt]
  X_{\text{MEAS,i}} &\sim \text{Normal}(X_{\text{TRUE,i}}, X_{\text{SE,i}}) \\[-1pt]
  X_{\text{TRUE,i}} &\sim \text{Normal}(\bar{X}_\text{TRUE}, \tau) \\[-1pt]
  \beta_0, \beta_1 &\sim \text{Normal}(0, 2) \\[-1pt]
  \sigma &\sim \text{Half Student t}(3, 0, 2) \\[-1pt]
  \bar{X}_\text{TRUE} &\sim \text{Normal}(0, 1) \\[-1pt]
  \tau &\sim \text{Half Student t}(3, 0, 2) \\
\end{aligned}
\end{equation*}

##

```{r}
tar_load(cont_coef_plot)
cont_coef_plot +
  theme(text = element_text(family = ""),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        plot.caption = element_text(size = 16),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        legend.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA")) +
  labs(y = "Coefficient Estimate", x = "Correlation Between True Ideal Point and\nMean Measurement Error Ideal Point",
         color = "Model Type")
```

# Skewed IRT Example

##

```{r}
tar_load(theta_dist_plot)
theta_dist_plot +
  theme(text = element_text(family = ""),
        plot.title = element_text(size = 22),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        plot.caption = element_text(size = 18),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"))
```

##

**Joint Model for Skewed Measurement Error**

\begin{equation*}
\begin{aligned}
  y_i &\sim \text{Normal}(\mu_i, \sigma) \\[-1pt]
  \mu_i &= \beta_0 + \beta_1 X_{\text{TRUE,i}} \\[-1pt]
  \color{teal}X_{\text{MEAS,i}} &\sim \color{teal}\text{Skew Normal}(X_{\text{TRUE,i}}, X_{\text{SE,i}}, \alpha_i) \\[-1pt]
  X_{\text{TRUE,i}} &\sim \text{Normal}(\bar{X}_\text{TRUE}, \tau_1) \\[-1pt]
  \color{teal}\bar{X}_\text{MEAS} - X_{\text{MEAS,i}} &\color{teal}\sim \text{Normal}(\alpha_i, \tau_2) \\[-1pt]
  \beta_0, \beta_1 &\sim \text{Normal}(0, 2) \\[-1pt]
  \sigma &\sim \text{Half Student t}(3, 0, 2) \\[-1pt]
  \bar{X}_\text{TRUE} &\sim \text{Normal}(0, 1) \\[-1pt]
  \tau_1 &\sim \text{Half Student t}(3, 0, 2) \\[-1pt]
  \color{teal}\alpha &\color{teal}\sim \text{Normal}(0, 1) \\[-1pt]
  \color{teal}\tau_2 &\color{teal}\sim \text{Half Student t}(3, 0, 2)
\end{aligned}
\end{equation*}

##

```{r}
tar_load(irt_coef_plot)
irt_coef_plot +
  theme(text = element_text(family = ""),
        plot.title = element_text(size = 22),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        plot.caption = element_text(size = 16),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"))
```

- Coefficient estimate attenuation of about 15% in realistic IRT example

# Thank you!