### 2023-08-16 AP_Retreat_2023-06-07/AP_Retreat_2023-06-07.Rmd
workshop presentation

---
title: "Going Beyond Ideal Point Points: Modeling Measurement Model Measurement Error"
subtitle: "AP Retreat"
author: "Bertrand Wilden"
date: \today
output: binb::metropolis
header-includes:
   - \usepackage{tikz}
   - \usetikzlibrary{positioning}
   - \usepackage{xcolor}
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache.lazy = FALSE)

pacman::p_load(
  "tidyverse",
  "here",
  "patchwork",
  "targets",
  "kableExtra",
  "tidybayes",
  "ggdist",
  "MetBrewer",
  "targets"
)

options(scipen = 1, digits = 3)
```

##

**Overview**

- Latent variables (e.g. political ideology) are often measured using statistical models

- When using these variables in downstream analysis, bias can occur if measurement uncertainty is ignored

- I develop a method for fixing this

##

**Measurement Models**

- Used to measure latent variables: *ideology, corruption, democracy, racial resentment*
- Example model: item response theory (IRT) ideal point model

**Theory-Testing Models**

- Testing a causal theory: *being an ideologically extreme candidate hurts one's reelection chances*
- Example model: regression with controls

##

**Bayesian Measurement Models**

- Output is a full posterior distribution of values for the latent variable according to their relative plausibility

- In practice, only a point estimate such as mean or median is used is subsequent analyses

##

```{r}
N <- 10000
tibble(x = c(rnorm(N, sd = 1),
             rnorm(N, sd = 2),
             sn::rsn(N, alpha = -4, omega = 2, xi = 1.5),
             sn::rst(N, xi = -2.75, omega = 2, alpha = 4, nu = 1)),
       y = c(rep("A\nNormal(0, 1)", N), 
             rep("B\nNormal(0, 2)", N), 
             rep("C\nSkew Normal(1.5, 2, -4)", N), 
             rep("D\nSkew t(-2.75, 2, 4, 1)", N))) |> 
  ggplot(aes(x = x)) +
  geom_histogram(color = "white", fill = "darkcyan", alpha = .5, bins = 50) +
  geom_vline(xintercept = 0, linetype = "twodash", color = "black", linewidth = .7) +
  # geom_text(aes(label = y), x = -4, y = 0) +
  facet_wrap(~ y, nrow = 2) +
  xlim(-5, 5) +
  theme_ggdist() +
  labs(y = "", x = "", caption = "Four different measurement model posterior distributions with mean zero") +
  theme(axis.line.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.x = element_text(size = 18),
        plot.title = element_text(hjust = .5, size = 24),
        text = element_text(family = ""),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        plot.caption = element_text(size = 18),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"))
```

##

**Three Problems With Ignoring Measurement Model Uncertainty**

::: incremental
  1. Uncertainty is not properly propagated in theory-testing models using point estimates
    - ***Coefficient uncertainty is underestimated***
  2. There is too much random noise in the point estimates
    - ***Coefficients are biased towards zero***
  3. Possibility of backdoor confounding through the measurement process
    - ***Coefficients are biased in unpredictable ways***
:::

##

\centering

\textbf{Solution: Joint Bayesian Measurement Error Model}

\raggedright

\begin{equation*}
\begin{aligned}
   y_i  &\sim f(\theta_i) \\
   y^{*}_i &\sim g(\theta_i)
\end{aligned}
\end{equation*}

- $y$, outcome in theory-testing model
- $\theta$, latent variable
- $y^*$, training data in measurement model
- $f(\cdot)$, theory-testing model
- $g(\cdot)$, measurement model

##

**Simplifications**

::: incremental
- Applied researchers know a lot about $f(\cdot)$, but not $g(\cdot)$
- The real $g(\cdot)$ can take a lot of computation time
- My solution is to use an approximation of $g(\cdot)$ that fits the measurement error posterior distribution
:::

##

\centering

\textbf{Joint IRT Ideal Point Measurement Error Model}

\raggedright

\begin{equation*}
\begin{aligned}
  y_i &\sim \text{Normal}(\beta_0 + \beta_1 \textcolor{teal}{\theta_i}, \sigma^2) \\
  y^*_{ij} &\sim \text{Bernoulli}[\Phi(\gamma_j\textcolor{teal}{\theta_i} + \xi_i)] \\
\end{aligned}
\end{equation*}

Is simplified as:

\begin{equation*}
\begin{aligned}
  y_i &\sim \text{Normal}(\beta_0 + \beta_1 \textcolor{teal}{\theta_i}, \sigma^2) \\
  \bar{\theta}_i &\sim \text{Normal}(\textcolor{teal}{\theta_i}, \sigma^2_{\theta [i]})
\end{aligned}
\end{equation*}

- $\bar{\theta}_i$, posterior mean from measurement model
- $\sigma^2_{\theta [i]}$, posterior variance from measurement model

# Measurement Error and Attenuation Bias

##

**Measurement Model**

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{Large}
\begin{center}
\tikz{
    \node (theta_s) {$\bar{\theta}$};
    \node (theta_label) [right = 0 of theta_s] {$\textcolor{teal}{\text{\normalsize Estimated Ideal Point}}$};
    \node[hidden, dashed] (theta) [above = 1.5 of theta_s] {$\theta$};
    \node (theta_label) [above = .1 of theta] {$\textcolor{teal}{\text{\normalsize True Ideal Point}}$};
    \node (e_theta) [left = 1.5 of theta_s] {$e^\theta$};
    \node (e_theta_label) [left = 0 of e_theta] {$\textcolor{teal}{\text{\normalsize Measurement Error}}$};
    \node (p) [below = 1.5 of e_theta] {$P$};
    \node (p_label) [below = 0 of p] {$\textcolor{teal}{\text{\normalsize Participation}}$};

    \path (e_theta) edge (theta_s);
    \path[dashed] (theta) edge (theta_s);
    \path (p) edge [bend left=0] (e_theta);
    \path[dashed] (theta) edge [bend right=30] (e_theta);
}
\end{center}
\end{Large}

##

**Theory-testing Model**

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{Large}
\begin{center}
\tikz{
    \node (theta_s) {$\bar{\theta}$};
    \node[hidden, dashed] (theta) [above = 1.5 of theta_s] {$\theta$};
    \node (e_theta) [left = 1.5 of theta_s] {$e^\theta$};
    \node (p) [below = 1.5 of e_theta] {$P$};
    \node (y) [right = 1.5 of theta_s] {$Y$};
    \node (y_label) [above = 0 of y] {$\textcolor{teal}{\text{\normalsize Outcome}}$};

    \path (e_theta) edge (theta_s);
    \path (theta_s) edge (y);
    \path[dashed] (theta) edge (theta_s);
    \path (p) edge [bend left=0] (e_theta);
    \path[dashed] (theta) edge [bend right=30] (e_theta);
}
\end{center}
\end{Large}

##

**Simulation Study**

:::incremental
1. Generate latent variable distributions from true $\theta_i$
2. Generate outcome $y_i$ as a function of true $\theta_i$
3. Use maximum likelihood to estimate distributional parameters (mean, variance, skew) from step 1 for use in simplified measurement model
4. Compare parameter recovery using a mean-only model and a joint measurement error model
:::

##

\textbf{Mean-Only Model}

\begin{equation*}
\begin{aligned}
  y_i &\sim \text{Normal}(\mu_i, \sigma^2) \\
  \mu_i &= \beta_0 + \beta_1 \textcolor{brown}{\bar{\theta}_i} \\
  \beta_0, \beta_1 &\sim \text{Normal}(0, 2) \\
  \sigma &\sim \text{Half Student t}(3, 0, 2) \\
\end{aligned}
\end{equation*}

##

\textbf{Joint Measurement Error Model}

\begin{equation*}
\begin{aligned}
  y_i &\sim \text{Normal}(\mu_i, \sigma^2) \\
  \mu_i &= \beta_0 + \beta_1 \textcolor{teal}{\theta_i} \\
  \textcolor{teal}{\bar{\theta}_i} &\textcolor{teal}{\sim \text{Normal}(\theta_i, \sigma^2_{\theta [i]})} \\
  \textcolor{teal}{\theta_i} &\textcolor{teal}{\sim \text{Normal}(\pi, \tau)} \\
  \beta_0, \beta_1 &\sim \text{Normal}(0, 2) \\
  \sigma &\sim \text{Half Student t}(3, 0, 2) \\
  \textcolor{teal}{\tau} &\textcolor{teal}{\sim \text{Half Student t}(3, 0, 2)} \\
  \textcolor{teal}{\pi} &\textcolor{teal}{\sim \text{Normal}(0, 1)}
\end{aligned}
\end{equation*}

##

```{r}
tar_load(me_cont)
tar_load(no_me_cont)
true = 1

coefs = rbind(me_cont, no_me_cont) |> 
    mutate(model_name = case_when(model_name == "Mean Values Only" ~ "Mean Only Model",
                                  model_name == "Measurement Error" ~ "Measurement Error Model"))
```

```{r}
coefs |> 
  filter(model_name == "Mean Only Model") |> 
    ggplot(aes(x = corr_x, y = mean, color = model_name)) +
    geom_smooth(n = 8, fill = "gray", level = 0.89) +
    geom_pointrange(aes(ymin = `5.5%`, ymax = `94.5%`), size = .6) +
    geom_hline(yintercept = true, linetype = "dashed") +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_ggdist() +
    scale_y_continuous(breaks = c(0, .5, 1), limits = c(-.1, 1.4)) +
    scale_x_reverse(limits = c(1, 0)) +
    labs(y = "Coefficient Estimate", x = "Correlation Between True Ideal Point and\nMean Measurement Error Ideal Point",
         color = "Model Type", caption = paste("True Coefficient Value =", true)) +
  theme(text = element_text(family = ""),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        plot.caption = element_text(size = 16),
        legend.position = "top",
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        legend.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"))+
  scale_color_manual(values = MetPalettes$Isfahan1[[1]][2])
```

##

```{r}
coefs |> 
    ggplot(aes(x = corr_x, y = mean, color = model_name)) +
    geom_smooth(n = 8, fill = "gray", level = 0.89) +
    geom_pointrange(aes(ymin = `5.5%`, ymax = `94.5%`), size = .6) +
    geom_hline(yintercept = true, linetype = "dashed") +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_ggdist() +
    scale_color_met_d("Isfahan1") +
    scale_x_reverse(limits = c(1, 0)) +
    labs(y = "Coefficient Estimate", x = "Correlation Between True Ideal Point and\nMean Measurement Error Ideal Point",
         color = "Model Type", caption = paste("True Coefficient Value =", true)) +
  theme(text = element_text(family = ""),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        plot.caption = element_text(size = 16),
        legend.position = "top",
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        legend.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"))
```

# Measurement Error and Confounding Bias

##

**Theory-testing Model**

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{Large}
\begin{center}
\tikz{
    \node (theta_s) {$\bar{\theta}$};
    \node[hidden, dashed] (theta) [above = 1.5 of theta_s] {$\theta$};
    \node (e_theta) [left = 1.5 of theta_s] {$e^\theta$};
    \node (p) [below = 1.5 of e_theta] {$P$};
    \node (y) [right = 1.5 of theta_s] {$Y$};
    \node[hidden, dashed] (u) [below = 1.5 of theta_s] {$U$};
    \node (u_label) [below = 0 of u] {$\textcolor{teal}{\text{\normalsize Confound}}$};

    \path (e_theta) edge (theta_s);
    \path (theta_s) edge (y);
    \path[dashed] (theta) edge (theta_s);
    \path (p) edge [bend left=0] (e_theta);
    \path[dashed] (theta) edge [bend right=30] (e_theta);
    \path[dashed] (u) edge [bend left = 30] (e_theta);
    \path[dashed] (u) edge [bend right = 30](y);
}
\end{center}
\end{Large}

##

\textbf{Skew-Normal Joint Measurement Error Model}

\begin{equation*}
\begin{aligned}
  y_i &\sim \text{Normal}(\mu_i, \sigma^2) \\
  \mu_i &= \beta_0 + \beta_1 \theta_i \\
  \textcolor{teal}{\bar{\theta}_i} &\textcolor{teal}{\sim \text{Skew Normal}(\theta_i, \omega_{\theta[i]}, \alpha_{\theta[i]})} \\
  \theta_i &\sim \text{Normal}(\pi, \tau) \\
  \beta_0, \beta_1 &\sim \text{Normal}(0, 2) \\
  \sigma &\sim \text{Half Student t}(3, 0, 2) \\
  \tau &\sim \text{Half Student t}(3, 0, 2) \\
  \pi &\sim \text{Normal}(0, 1)
\end{aligned}
\end{equation*}

##

**Skew-Normal Measurement Error Model Comparison**

```{r}
tar_load(bias_coef_plot)
bias_coef_plot +
  theme(text = element_text(family = ""),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20),
        plot.caption = element_text(size = 18),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        legend.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"))
```

# Case Study: Ideological Extremism and Electoral Success

##

**Research Question:** are ideologically extreme US House incumbents punished electorally?

- DV: General election vote share
- IV: Ideology estimated from previous Congress roll-call votes
- Data from 1990 to 2016

##

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{Large}
\begin{center}
\tikz{
    \node (theta_s) {$\bar{\theta}$};
    \node[hidden, dashed] (theta) [above = 1.5 of theta_s] {$\theta$};
    \node (e_theta) [left = 1.5 of theta_s] {$e^\theta$};
    \node (y) [right = 1.5 of theta_s] {$Y$};
    \node (y_label) [above = 0 of y] {$\textcolor{teal}{\text{\normalsize Incum. Vote \%}}$};
    \node[hidden, dashed] (u) [below = 1.5 of theta_s] {$U$};
    \node[hidden, dashed] (Theta) [above = 1.5 of y] {$\Theta$};
    \node (Theta_label) [above = 0 of Theta] {$\textcolor{teal}{\text{\normalsize District Ideology}}$};
    \node (r) [right = 1.5 of y] {$R$};
    \node (r_label) [below = 0 of r] {$\textcolor{teal}{\text{\normalsize Repub. Pres. Vote \%}}$};
    \node[hidden, dashed] (e_Theta) [right = 1.5 of r] {$e^\Theta$};

    \path (e_theta) edge (theta_s);
    \path (theta_s) edge (y);
    \path[dashed] (theta) edge (theta_s);
    \path[dashed] (theta) edge [bend right=30] (e_theta);
    \path[dashed] (u) edge [bend left = 30] (e_theta);
    \path[dashed] (u) edge [bend right = 30](y);
    \path[dashed] (Theta) edge (r);
    \path (r) edge (y);
    \path (e_Theta) edge (r);
    \path[dashed] (Theta) edge (theta_s);
}
\end{center}
\end{Large}

##

**Effect of Incumbent Ideology on General Election Vote Share**

```{r}
tar_load(reelection_coef_plot)
reelection_coef_plot +
  theme(text = element_text(family = ""),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20),
        plot.caption = element_text(size = 18),
        strip.text = element_text(size = 18),
        plot.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        panel.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"),
        legend.background = element_rect(fill = "#FAFAFA", color = "#FAFAFA"))
```

# Concluding Thoughts

##

- Uncertainty estimation matters
  - Better to use measurement models that provide uncertainty estimates than those that don't (eg NOMINATE)
  - Better to use state-of-the-art posterior sampling methods such as Hamiltonian Monte Carlo, rather than traditional Gibbs samplers for measurement models

\tikzset{
    > = stealth,
    every node/.append style = {
        draw = none
    },
    every path/.append style = {
        arrows = ->
    },
    hidden/.style = {
        draw = black,
        shape = circle,
        inner sep = 1pt
    }
}

\begin{normalsize}
\begin{center}
\tikz{
    \node (theta_s) {$\bar{\theta}$};
    \node[hidden, dashed] (theta) [above = 1 of theta_s] {$\theta$};
    \node (e_theta) [left = 1.5 of theta_s] {$e^\theta$};
    \node (e_theta_label) [below = 0 of e_theta] {$\textcolor{teal}{\text{\normalsize Observed Error}}$};
    \node[hidden, dashed] (eo_theta) [right = 1.5 of theta_s] {$e^\theta$};
    \node (eo_theta_label) [below = 0 of eo_theta] {$\textcolor{teal}{\text{\normalsize Unobserved Error}}$};

    \path (e_theta) edge (theta_s);
    \path[dashed] (theta) edge (theta_s);
    \path[dashed] (eo_theta) edge (theta_s);
}
\end{center}
\end{normalsize}

# Thank you!